% Copyright (C) 2008, 2009, 2010, 2011, 2012, 2014 Bert Burgemeister
%
% Permission is granted to copy, distribute and/or modify this
% document under the terms of the GNU Free Documentation License,
% Version 1.2; with no Invariant Sections, no Front-Cover Texts and
% no Back-Cover Texts. For details see file COPYING.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Arrays} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{section:Arrays}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Predicates} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{LIST}{1cm}

  \IT{\arrGOO{%
      (\FU*{ARRAYP} \VAR{ foo})\\
      (\FU*{VECTORP} \VAR{ foo})\\
      (\FU*{SIMPLE-VECTOR-P} \VAR{ foo})\\
      (\FU*{BIT-VECTOR-P} \VAR{ foo})\\
      (\FU*{SIMPLE-BIT-VECTOR-P} \VAR{ foo})}{.}}
  {
  \retval{\T} if \VAR{foo} is of indicated type.
  }

  \IT{\arrGOO{(\FU*{ADJUSTABLE-ARRAY-P} \VAR{ array})\\
      (\FU*{ARRAY-HAS-FILL-POINTER-P} \VAR{ array})}{.}}
  {
  \retval{\T} if \VAR{array} is adjustable/has a fill pointer,
  respectively. 
  }

  \IT{(\FU*{ARRAY-IN-BOUNDS-P} \VAR{array} \Op{\VAR{subscripts}})}
  {
  Return \retval{\T} if \VAR{subscripts} are in \VAR{array}'s bounds.
  }
  % No default subscripts in standard.

\end{LIST}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Array Functions} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{LIST}{1cm}

  \IT{(\xorGOO{%
      \FU*{MAKE-ARRAY}\VAR{ dimension-sizes }
      \OP{\kwd{:adjustable} \VAR{ bool}\DF{\NIL}}\\
      \FU*{ADJUST-ARRAY } \DES{\VAR{array}}
      \VAR{ dimension-sizes}}{\}}
    \orGOO{\kwd{:element-type} \VAR{ type}\DF{\T}\\
      \kwd{:fill-pointer } \Goo{\VAR{num}\XOR\VAR{bool}}\DF{\NIL}\\
      \xorGOO{\kwd{:initial-element} \VAR{ obj}\\
        \kwd{:initial-contents} \VAR{ tree-or-array}\\
        \kwd{:displaced-to } \VAR{array}\DF{\NIL}\text{ }
        \Op{\kwd{:displaced-index-offset } \VAR{i}\DF{\LIT{0}}}}{.}}{\}})}
  { 
  Return fresh, or readjust, respectively, \retval{vector} or \retval{array}. 
  }

  \IT{(\FU*{AREF} \VAR{array} \OP{\VAR{subscripts}})}
  {
  Return
  \retval{array element} pointed to by \VAR{subscripts}. \kwd{setf}able.
  }
  % No subscripts means sole element of rank zero array.

  \IT{(\FU*{ROW-MAJOR-AREF} \VAR{array} \VAR{i})}
  {
  Return \retval{\VAR{i}th element} of \VAR{array} in row-major
  order. \kwd{setf}able.
  }

  \IT{(\FU*{ARRAY-ROW-MAJOR-INDEX} \VAR{array} \Op{\VAR{subscripts}})}
  {
  \retval{Index} in row-major order of the element denoted by \VAR{subscripts}.
  }

  \IT{(\FU*{ARRAY-DIMENSIONS} \VAR{array})}
  {
  \retval{List} containing the lengths of \VAR{array}'s dimensions. 
  }

  \IT{(\FU*{ARRAY-DIMENSION} \VAR{array} \VAR{i})}
  {
  \retval{Length of \VAR{i}th dimension} of \VAR{array}.
  }

  \IT{(\FU*{ARRAY-TOTAL-SIZE} \VAR{array})}
  {
    \retval{Number of elements} in \VAR{array}.
  }

  \IT{(\FU*{ARRAY-RANK} \VAR{array})}
  {
    \retval{Number of dimensions} of \VAR{array}.
  }

  \IT{(\FU*{ARRAY-DISPLACEMENT} \VAR{array})}
  {
  \retval{Target array} and \retvalii{offset}.
  }

  \IT{\arrGOO{(\FU*{BIT} \VAR{ bit-array } \Op{\VAR{subscripts}})\\
      (\FU*{SBIT} \VAR{ simple-bit-array } \Op{\VAR{subscripts}})}{.}}
  {
  Return \retval{element} of \VAR{bit-array} or of
  \VAR{simple-bit-array}. \kwd{setf}\-able. 
  }

  \IT{(\FU*{BIT-NOT} \DES{\VAR{bit-array}} \Op{\DES{\VAR{result-bit-array}}\DF{\NIL}})}
  {
  Return \retval{result} of bitwise negation of \VAR{bit-array}. If \VAR{result-bit-array} is \T, put
  result in \VAR{bit-array}; if it is \NIL, make a new array for
  result. 
  }

  \IT{(\xorGOO{%
      \FU*{BIT-EQV}\\
      \FU*{BIT-AND}\\
      \FU*{BIT-ANDC1}\\
      \FU*{BIT-ANDC2}\\
      \FU*{BIT-NAND}\\
      \FU*{BIT-IOR}\\
      \FU*{BIT-ORC1}\\
      \FU*{BIT-ORC2}\\
      \FU*{BIT-XOR}\\
      \FU*{BIT-NOR}}{\}} \DES{\VAR{bit-array-a}} \VAR{bit-array-b}
    \Op{\DES{\VAR{result-bit-array}}\DF{\NIL}})}
  {
  Return \retval{result} of bitwise logical operations
  (cf.\ operations of \FU{boole}, page \pageref{section:Logic Functions}) on
  \VAR{bit-array-a} and  \VAR{bit-array-b}.  If \VAR{result-bit-array} is \T, put result in
  \VAR{bit-array-a}; if it is \NIL, make a new array for
  result. 
  }

  \IT{\CNS*{ARRAY-RANK-LIMIT}}
  {
  Upper bound of array rank; $\geq 8$.
  }

  \IT{\CNS*{ARRAY-DIMENSION-LIMIT}}
  {
  Upper bound of an array dimension; $\geq 1024$.
  }

  \IT{\CNS*{ARRAY-TOTAL-SIZE-LIMIT}}
  {
  Upper bound of array size; $\geq 1024$.
  }

\end{LIST}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Vector Functions} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Vectors can as well be manipulated by sequence functions; see
section~\ref{section:Sequences}. 

\begin{LIST}{1cm}
  
  \IT{(\FU*{VECTOR} \OPn{\VAR{foo}})}
  {
    Return fresh \retval{simple vector of \VAR{foo}s}.
  }
  % No default foo in standard.

  \IT{(\FU*{SVREF} \VAR{vector} \VAR{i})}
  {
    \retval{Element \VAR{i}} of simple \VAR{vector}. \kwd{setf}able.
  }

  \IT{(\FU*{VECTOR-PUSH} \VAR{foo} \DES{\VAR{vector}})}
  {
  Return \retval{\NIL} if \VAR{vector}'s fill pointer equals size of
  \VAR{vector}. Otherwise replace element of \VAR{vector} pointed to
  by \retval{fill pointer} with \VAR{foo}; then increment fill
  pointer. 
  }

  \IT{(\FU*{VECTOR-PUSH-EXTEND} \VAR{foo} \DES{\VAR{vector}}
    \Op{\VAR{num}})}
  {
  Replace element of \VAR{vector} pointed to by \retval{fill pointer} with
  \VAR{foo}, then increment fill pointer. Extend \VAR{vector}'s size by
  $\ge \VAR{num}$ if necessary.
  }

  \IT{(\FU*{VECTOR-POP} \DES{\VAR{vector}})}
  {
  Return \retval{element of \VAR{vector}} its fillpointer points to
  after decrementation.
  }

  \IT{(\FU*{FILL-POINTER} \VAR{vector})}
  {
  \retval{Fill pointer} of \VAR{vector}. \kwd{setf}able.
  }

\end{LIST}



%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "clqr"
%%% End: 
