%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Symbols and Packages}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{LIST}{1cm}

  \IT{\arrGOO{(\FU*{SYMBOLP} \VAR{foo})\\
      (\FU*{PACKAGEP} \VAR{foo})}{\}}}
  Return \retval{\T} if \VAR{foo} is of type \kwd{symbol} or
  \kwd{package}, resp.

  \IT{(\FU*{BOUNDP} \VAR{symbol})}
  \retval{\T} if \VAR{symbol} is bound.

  \IT{(\FU*{KEYWORDP} \VAR{foo})}
  \retval{\T} if \VAR{foo} is a keyword.

  \IT{(\kwd*{T})}
  Truth, or \V{*terminal-io*}.

  \IT{\KWD*{:}\VAR{foo}} 
  Keyword, evaluates to \retval{:\VAR{foo}}

  \IT{\Goo{(\CNS*{NIL})\XOR{\kwd*{()}}}} 
  Falsehood, the empty list, \V{*standard-input*},
  \V{*standard-output*}, or the global environment.

  \IT{\V*{*package*}} The current package.

  \IT{(\FU*{SYMBOL-PACKAGE '}\VAR{foo})} 
  Return the \retval{package \VAR{foo} came from}.

  \IT{(\GFU*{DOCUMENTATION} \kwd{'}\VAR{foo}
    \Goo{\kwd{variable}\XOR\kwd{function}\XOR
      \kwd*{compiler-macro}\XOR
      \kwd{method-combination}\XOR
      \T\XOR\kwd{structure}
      \XOR\kwd{type}\XOR\kwd{setf}\XOR\T})}
  Return \retval{documentation string} of \VAR{foo} of given
  type. \kwd{setf}able. 

  \IT{(\kwd*{DEFPACKAGE}\VAR{ foo} 
    \orGOO{\OPn{(\kwd{:use }\OPn{\VAR{bar}})}\\
      \OPn{(\kwd{:nicknames }\OPn{\VAR{nick}})}\\
      (\kwd{:documentation }\VAR{doc})\\
      \OPn{(\kwd{:export }\OPn{\VAR{baz}})}\\
      \OPn{(\kwd{:import-from }\VAR{p} \OPn{\VAR{n}})}\\
      \OPn{(\kwd{:intern }\OPn{\VAR{n}})}\\
      \OPn{(\kwd{:shadow}\XOR\Goo{\kwd{:shadowing-import-from}\VAR{ p}\OPn{\VAR{ jnk}}})}\\
      (\kwd{:size }\VAR{int})\\
    }{\}})} 
  Return \retval{defined package \VAR{foo}} which inherits symbols
  exported by packages \VAR{bar} excluding \VAR{jnk}s and some
  symbols. \VAR{n}s from package \VAR{p}; exports \VAR{baz}es. 
  
  \IT{(\kwd*{IN-PACKAGE }\VAR{foo})} 
  Make \retval{package \VAR{foo}} current.

  \IT{(\FU*{USE-PACKAGE} (\OPn{\VAR{foo}})
    \Op{\VAR{package}}\DF{\kwd{*package*}})}  
  Make exported symbols of packages \VAR{foo} available in
  \VAR{package}. Return \retval{\T}. 

  \IT{(\FU*{UNUSE-PACKAGE} (\OPn{\VAR{foo}})
    \Op{\VAR{package}}\DF{\kwd{*package*}})}
  Remove packages \VAR{foo} from use-list of \VAR{package}.

  \IT{(\FU*{DELETE-PACKAGE} \VAR{package})} Delete \VAR{package}. Return
  \retval{\T} if successful.

  \IT{(\FU*{FIND-PACKAGE} \Goo{\VAR{name}\XOR\VAR{package}})} Return
  \retval{package object} with \VAR{name} (case-sensitive).

  \IT{(\FU*{LIST-ALL-PACKAGES})} \retval{List of registered packages}.

  \IT{(\FU*{MAKE-PACKAGE} \VAR{name}
    \orGOO{\kwd{:nicknames }(\OPn{\VAR{nick}})\DF{\NIL}\\
      \kwd{:use }(\OPn{package})}{\}})}
  Create \retval{package \VAR{name}}.

  \IT{(\xorGOO{%
      \xorGOO{%
        \FU*{DO-ALL-SYMBOLS}\\
        \FU*{DO-SYMBOLS}}{\}} 
      (\VAR{var} \OP{\VAR{package} \Op{\VAR{result}}})\\
    \FU*{DO-EXTERNAL-SYMBOLS} (\VAR{var} \Op{\VAR{result}})}{\}}
     \OPn{(\kwd{declare} \OPn{\VAR{decl}})} 
     \orGOO{\OPn{tag}\\
       \OPn{form}}{\}})}
  Evaluate \VAR{form}s with \VAR{var} successively bound to all
  symbols from all registered packages, to all symbols from
  \VAR{package}, or to all external symbols from \VAR{package}, resp. Return
  \retval{value of \VAR{result}}.

  \IT{(\FU*{EXPORT} \VAR{symbols} \Op{\VAR{package}\DF{\kwd{*package*}}})}
  Make \VAR{symbols} external to \VAR{package}. Return \retval{\T}.

  \IT{(\FU*{UNEXPORT} \VAR{symbols} \VAR{package})} 
  Revert \VAR{symbols} to internal status. Return \retval{\T}.

  \IT{(\FU*{PACKAGE-NAME} \VAR{package})} \retval{Name of \VAR{package}}.

  \IT{(\FU*{PACKAGE-NICKNAMES} \VAR{package})} 
  \retval{List of nicknames} of \VAR{package}.

  \IT{(\FU*{PACKAGE-SHADOWING-SYMBOLS} \VAR{package})} 
  \retval{List of shadowing symbols} of \VAR{package}.

  \IT{(\FU*{PACKAGE-USE-LIST} \VAR{package})}
  \retval{List of other packeges} used by \VAR{package}.

  \IT{(\FU*{PACKAGE-USED-BY-LIST} \VAR{package})}
  \retval{List of other packages} using \VAR{package}.

  \IT{(\FU*{RENAME-PACKAGE} \VAR{package} \Op{(\OPn{nick})\DF{\NIL}})}
  Rename \VAR{package}. Return \retval{renamed package}.

  \IT{(\SF*{WITH-PACKAGE-ITERATOR} (\VAR{name} \RP{\VAR{package}}
    \Op{\kwd{:internal}\XOR\kwd{:external}\XOR\kwd{:inherited}})
    \OPn{(\kwd{declare} \OPn{\VAR{decl}})} \OPn{\VAR{form}})}
  Return \retval{values of \VAR{form}s}. In the body, successive
  invocations of  (\VAR{name}) return: \T\ if a symbol is returned;
  a symbol from a \VAR{package}; accessibility
  (\kwd{:internal}, \kwd{:external}, or \kwd{:inherited}); and the
  package the symbol belongs to.

  \IT{(\xorGOO{\FU*{INTERN}\\
      \FU*{FIND-SYMBOL}}{\} \VAR{name} \Op{\VAR{package}}})}
  Intern or find, resp., \retval{symbol \VAR{name}} in
  \VAR{package}. Second return value is one of
  \retvalii{\kwd{:internal}}, \retvalii{\kwd{:external}}, or
  \retvalii{\kwd{:inherited}}.

  \IT{(\FU*{SHADOW} \RP{\VAR{symbol}}
    \Op{\VAR{package}\DF{\kwd{*package*}}})}
  Add \VAR{symbol}s to shadowed list of \VAR{package}. Return \retval{\T}.
  
  \IT{(\xorGOO{\FU*{IMPORT}\\
      \FU*{SHADOWING-IMPORT}}{\}}\RP{\VAR{symbol}}
    \Op{\VAR{package}\DF{\kwd{*package*}}})}
  Make \VAR{symbol}s internal  to \VAR{package}. Return \retval{\T}. In
  case of a name conflict, signal \kwd{package-error}, or shadow
  the old symbol, resp.

  \IT{(\kwd*{UNINTERN} \VAR{symbol} \Op{\VAR{package}})} 
  Remove \VAR{symbol} from \VAR{package}, return \retval{\T} on success.

  \IT{(\FU*{FIND-ALL-SYMBOLS} \VAR{string})}
  Return \retval{list of symbols} with name \VAR{string} from all
  registered packages.

\end{LIST}


